\section*{Referee Report R2}

\textbf{Recommendation:} minor revision\\
\textbf{Confidence:} high\\
\textbf{Manuscript:} \texttt{GPD/review-input/agent-economy-fraud-arxiv.md}\\
\textbf{SHA-256:} \texttt{3246d71037387a1d5bafc76becd9cdb40b7cac1179f8c60f7a2867a8f79c6f5b}

\subsection*{Summary}

The revision resolves the blocking issues from the first review round. The manuscript no longer claims necessary failure, absolute impossibility, production readiness, or a validated 6--12 month exploitation forecast. It now reads as an early measurement and detection-framework preprint with explicit empirical boundaries.

The strongest evidence remains the synthetic validation and attack-injection evaluation. The weak cleaned real-data AUC, noisy negative labels, threshold re-optimization, inactive Cross-Platform signal, and Chain 7 collective-detection gap are now visible in the abstract, evaluation, limitations, and conclusion.

\subsection*{Resolved Major Issues}

\begin{itemize}
\item \textbf{Claim scope:} resolved. The central claim is narrowed to human-invariant-only detectors under tested A2A scenarios.
\item \textbf{Literature positioning:} materially improved for the narrowed scope.
\item \textbf{Claim-test mapping:} resolved by the new claim-evidence map in the evaluation section.
\item \textbf{Empirical caveats:} resolved. Fixed-threshold recall, re-optimized recall, cleaned AUC, label noise, and inactive CP are foregrounded.
\item \textbf{Venue/readiness framing:} resolved. The paper is now framed as early measurement and framework work.
\end{itemize}

\subsection*{Minor Remaining Issue}

Archive platform documentation before external submission. OpenClaw and Moltbook are current platform sources. For long-term reproducibility, the authors should archive or snapshot the exact source versions used for the threat model.

\subsection*{Recommendation}

Minor revision. The remaining issue is archival and reproducibility-related rather than conceptual. The manuscript's revised claim scope is proportionate to the evidence.
