\begin{abstract}
Financial fraud detection systems are built on nine behavioral invariants
that describe properties exclusive to human actors: velocity limits,
biometric authentication, device fingerprinting, geographic constraints,
cognitive and energy constraints, identity persistence, behavioral
stability, computational limits, and bounded rationality. We show that
agent-to-agent (A2A) commerce stresses these invariants in systematic
ways: some current human-assumption signals become unreliable for
legitimate agent activity, while other agent-native signals become
available. Using the OpenClaw agent gateway and Moltbook agent social
platform as case studies, we derive an eight-chain taxonomy of A2A
fraud attack vectors grounded in platform documentation. We design a
five-signal agent-invariant detection framework and evaluate it on
81,904 Base-chain USDC transactions from 665 ERC-8004 registered
agents, followed by a 93,579-transaction real corpus augmented with
6,050 injected attacks. On real data the framework achieves recall of
95.4\% on the raw, contaminated-negative label set after threshold
re-optimization (81.1\% after label cleaning); at the fixed
synthetic-calibrated threshold, raw-label recall is 61.4\%, and composite
ROC-AUC is only 0.515 on cleaned labels. These figures show useful recall
under a noisy-label operating point, not mature ranking performance.
Against a labeled mixed dataset with injected attack patterns, the
framework detects 7 of~8 attack chains at 100\% per-chain recall. The
eighth chain---coordinated swarm attacks---reveals an architectural gap:
per-address scoring is blind to coordinated multi-agent attacks where
individual agents have minimal on-chain footprint, motivating a
next-generation collective detection architecture. We provide an
open-source reference implementation (Apache~2.0) and separate
framework-validated recommendations from broader policy proposals.
We position this work as an empirical, detection-side complement to the
emerging agentic-commerce-security literature: as of mid-2026, agent
impersonation and agent-wallet theft are already documented, while
systematic A2A-commerce fraud exploiting the specific invariant blind
spots modeled here remains undocumented and is the threat model's
decisive future test.

\medskip
\noindent\textbf{Keywords:} agent-to-agent commerce, fraud detection,
behavioral invariants, multi-agent systems, blockchain analytics,
financial security, ERC-8004
\end{abstract}
