\section{Reproducibility}\label{app:reproducibility}

This appendix provides all information necessary to reproduce the experimental
results reported in this paper.

%----------------------------------------------------------------------
\subsection{Software Package}

The reference implementation is released as \texttt{a2a-fincrime-detection}
version 0.1.0, installable from PyPI under the package name
\texttt{a2a-detection}.

\begin{itemize}[leftmargin=*]
  \item \textbf{Language:} Python $\geq$ 3.10
  \item \textbf{License:} Apache 2.0
  \item \textbf{Dependencies:} \texttt{httpx}, \texttt{pandas},
    \texttt{pyarrow}, \texttt{web3}, \texttt{python-dotenv},
    \texttt{numpy}, \texttt{scikit-learn}
\end{itemize}

%----------------------------------------------------------------------
\subsection{Installation}

Install from PyPI:

\begin{lstlisting}[language=bash]
pip install a2a-detection
\end{lstlisting}

Alternatively, install from source:

\begin{lstlisting}[language=bash]
git clone https://github.com/<repo>/a2a-fincrime-detection.git
cd a2a-fincrime-detection
pip install -e .
\end{lstlisting}

The source distribution uses \texttt{pyproject.toml} for build configuration.

%----------------------------------------------------------------------
\subsection{Reproducing Experiments}

\paragraph{Step 1: Inject synthetic attacks.}
Generate the attack injection dataset with a fixed random seed for
deterministic output:

\begin{lstlisting}[language=bash]
python -m a2a_detection.scripts.inject_attacks --seed 42
\end{lstlisting}

\paragraph{Step 2: Run fraud detection validation.}
Execute the full detection pipeline on the labeled dataset and report
metrics:

\begin{lstlisting}[language=bash]
python -m a2a_detection.scripts.validate_fraud_detection
\end{lstlisting}

This command produces per-signal AUC scores, composite score distributions,
and the Phase~5 evaluation metrics reported in \Cref{sec:evaluation}.

\paragraph{Step 3: Run the test suite.}
Verify correctness of all components:

\begin{lstlisting}[language=bash]
python -m pytest src/a2a_detection/tests/ -v
\end{lstlisting}

All 47 tests should pass. The test suite covers signal computation,
threshold calibration, attack injection, and label loading.

%----------------------------------------------------------------------
\subsection{Data Availability}

The labeled dataset (see \Cref{app:dataset}) is released alongside the
package. After installation, data files are accessible via the package's
built-in data loader:

\begin{lstlisting}[language=python]
from a2a_detection.data import load_transactions, load_labels
txs = load_transactions()    # 81,904 rows
labels = load_labels()       # 1,734 labeled addresses
\end{lstlisting}

%----------------------------------------------------------------------
\subsection{Hardware Requirements}

\begin{itemize}[leftmargin=*]
  \item \textbf{CPU:} Standard consumer or server CPU; no GPU required.
  \item \textbf{Memory:} $<$ 4\,GB RAM for the full pipeline.
  \item \textbf{Inference latency:} 97\,ms per address on consumer hardware
    (Apple M2, single core).
  \item \textbf{Full evaluation:} Completes in under 60 seconds on the
    1{,}734-address dataset.
\end{itemize}
