% recommendations.tex — Section 7: Industry Recommendations

\section{Industry Recommendations}\label{sec:recommendations}

We translate our detection-framework findings into a prioritized
roadmap for financial institutions.  Each recommendation is scored on a
weighted composite of \emph{impact} (40\%), \emph{urgency} (30\%),
\emph{feasibility} (20\%), and \emph{cost} (10\%), then assigned to one
of four priority tiers.

%% -----------------------------------------------------------------
\subsection{Priority Matrix}\label{subsec:priorities}

\Cref{tab:priorities} summarizes the full roadmap.

\begin{table}[t]
  \centering
  \caption{Recommended deployment roadmap, ordered by composite
    priority score.  Cost estimates assume a mid-tier U.S.\ financial
    institution.}
  \label{tab:priorities}
  \small
  \begin{tabularx}{\textwidth}{@{}c l c X@{}}
    \toprule
    \textbf{ID} & \textbf{Recommendation} & \textbf{Tier / Horizon} & \textbf{Rationale} \\
    \midrule
    \multicolumn{4}{@{}l}{\textit{P0 --- Immediate (0--6 months, \$80K--\$200K)}} \\
    \addlinespace[2pt]
    R1 & Economic Rationality + Network Topology signals
       & P0 / 0--6\,mo
       & Highest individual AUC signals (NT\,=\,0.5990, ER\,=\,0.5152 on
         cleaned labels; see \cref{tab:signal_auc}).
         Deployable as scoring overlays on existing transaction-monitoring
         pipelines.  Estimated cost: \$80K. \\
    \addlinespace[2pt]
    R2 & Agent-platform metadata detection
       & P0 / 0--6\,mo
       & Parse session metadata for OpenClaw session-key patterns
         and ERC-8004 registry lookups.  Low engineering effort;
         estimated cost: \$20K. \\
    \midrule
    \multicolumn{4}{@{}l}{\textit{P1 --- Near-term (6--12 months, \$100K--\$500K)}} \\
    \addlinespace[2pt]
    R3 & Velocity-threshold adjustment
       & P1 / 6--12\,mo
       & Current thresholds ($>$1{,}000\,tx/hr) miss agent-speed
         activity.  Reduce to $>$100\,tx/hr for addresses flagged
         by R1--R2. \\
    \addlinespace[2pt]
    R4 & Cross-channel identity correlation
       & P1 / 6--12\,mo
       & Detect \texttt{identityLinks}-type aggregation across
         payment rails, flagging accounts whose identity graph
         expands faster than human norms. \\
    \addlinespace[2pt]
    R5 & Collective-behavior detection prototype
       & P1 / 6--12\,mo
       & Time-window clustering for \textsc{Chain\_7}:
         $\geq$10 new addresses transacting the same amount within
         a narrow time window. \\
    \midrule
    \multicolumn{4}{@{}l}{\textit{P2 --- Medium-term (12--24 months, \$500K--\$2M)}} \\
    \addlinespace[2pt]
    R6 & Biometric authentication mandate
       & P2 / 12--24\,mo
       & Require step-up biometric verification for agent-initiated
         transactions exceeding \$1{,}000. \\
    \addlinespace[2pt]
    R7 & Forensic data retention
       & P2 / 12--24\,mo
       & Preserve session transcripts, tool-call logs, and agent
         provenance metadata for 7~years to support post-incident
         investigation. \\
    \addlinespace[2pt]
    R8 & Agent KYC framework
       & P2 / 12--24\,mo
       & Extend KYC/AML obligations to the controlling human or
         legal entity behind agent wallets that transact above
         regulatory thresholds. \\
    \midrule
    \multicolumn{4}{@{}l}{\textit{P3 --- Long-term (24--36 months, \$2M+)}} \\
    \addlinespace[2pt]
    R9  & Multi-chain monitoring integration
        & P3 / 24--36\,mo
        & Extend detection to Ethereum L1, Arbitrum, Optimism, and
          Solana as agent activity spreads across chains. \\
    \addlinespace[2pt]
    R10 & Industry-standard agent-transaction labeling
        & P3 / 24--36\,mo
        & Contribute to or adopt a cross-industry standard for
          tagging agent-originated transactions in SWIFT/ISO~20022
          messages and on-chain metadata. \\
    \bottomrule
  \end{tabularx}
\end{table}

\paragraph{Sequencing rationale.}
P0 items deliver the highest risk-adjusted return: they leverage the two
strongest signals identified in \Cref{sec:evaluation} (Network Topology
AUC\,=\,0.5990, Economic Rationality AUC\,=\,0.5152 on cleaned labels) and require only
rule-layer additions to existing transaction-monitoring systems.
P1~items address velocity and coordination patterns that become
exploitable once agents operate at scale.  P2 and P3~items require
cross-functional investment (legal, compliance, infrastructure) and are
sequenced to align with anticipated regulatory timelines.

%% -----------------------------------------------------------------
\subsection{Privacy Compliance}\label{subsec:privacy}

All P0 recommendations are designed to comply with applicable data
protection and financial-crime regulations:

\begin{itemize}[nosep]
  \item \textbf{On-chain data.}  Blockchain transaction records are
    publicly available by design.  Analyzing agent addresses, transfer
    volumes, and temporal patterns on Base chain does not constitute
    processing of personal data under GDPR~\cite{gdpr2016} or
    CCPA~\cite{ccpa2018}, because on-chain addresses are pseudonymous
    and our detection pipeline does not attempt to link them to natural
    persons.
  \item \textbf{Agent-address analysis.}  The Economic Rationality and
    Network Topology signals operate on transaction-graph features
    (degree distribution, value entropy, timing regularity).  No
    personally identifiable information is ingested or produced.
  \item \textbf{Agent KYC (R8).}  When KYC obligations apply---\ie
    when agent-controlled wallets transact above thresholds set by the
    Bank Secrecy Act~\cite{bsa1970} and its implementing
    regulations---the obligation attaches to the \emph{controlling human
    or legal entity}, not to the agent itself.  This preserves the
    existing beneficial-ownership framework while closing the
    agent-intermediation gap.
  \item \textbf{Session-transcript retention (R7).}  Forensic retention
    of agent session logs must be implemented with appropriate access
    controls and data-minimization safeguards consistent with GDPR
    Article~5(1)(c) and GLBA Safeguards Rule requirements.
\end{itemize}

\noindent
In summary, the P0 and P1 tiers are implementable today without
regulatory friction; P2 items (R6--R8) will benefit from forthcoming
guidance on AI-agent financial activity that several jurisdictions are
actively developing.
